Managing variability in software systems

### 2021-06-04 Paper.tex
finish up

The essence of feature modeling is the splitting of features into distinct objects which have a certain relation to each other. There are several types of features mentioned in \cite{b8}, named \textit{mandatory, optional and alternative features}. Generally speaking common attributes of systems are described by mandatory features, while alternative or optional features define the variability in feature models.\cite{b9}\\

@@

\section{Variability patterns}
Variability can be introduced into a system via different mechanisms. These mechanisms strive to bind a given system to one out of many alternatives at different binding times. Binding times describe when exactly these mechanisms stop deffering the decision and make the choice for any of the variations, as illustrated in Figure 2.\cite{b2}

@@

Further explanation of these binding times can be found in the subsequent paragraphs:\\
\begin{itemize}
	\item \textit{Product architecture derivation.} The architecture of the product already binds to a variation. This is the lowest binding time and will influence all subsequent steps.\\
	\item \textit{Compilation.} When the source code is compiled there exist several options to bind to a variation by adjusting constants, macros or compiler directives.\\
	\item \textit{Linking.} This option does not exist for any technology, but for many programming languages the linking process can be manually changed. For example by switching the DLL of a module.\\
	\item \textit{Startup} A simple option to bind variants to the system is by configuration files that influence the features a software system exhibits.\\
	\item \textit{Adaptation to runtime environment} This typically refers to plugins of a software, which can add variation to a system and can either be provided by the original developer or even third party vendors. \\
\end{itemize}
Any mechanism to bind to none, one or many different variants tend to fall into different categories presented by \cite{b2} and \cite{b3}, and also have parallels to the feature types presented in the previous section:\\
\begin{itemize}
	\item \textit{Variant Entity.} There exist several options but only one may be bound to at any given time.\\
	\item \textit{Optional Entity.} An optional entity may or may not be included in the system.\\
	\item \textit{Multiple Coexisting Entities.} Similar to the variant entity but with the exception that multiple variants can be used.\\
\end{itemize}
These categories tend to be the same for the different binding times, but of course vary greatly in their implementation. While an optional entity during compilation may only be influenced by a compiler flag, it might be a whole plugin from a third party vendor when adapting a given runtime environment. An in depth analysis of these mechanisms can be found in \cite{b2}.

>>>>

The essence of feature modeling is the splitting of features into distinct objects which have a certain relation to each other. There are several types of features mentioned in \cite{b8}, named \textit{mandatory, optional and alternative features}. Generally speaking common attributes of systems are described by mandatory features, while alternative or optional features define the variability in feature models. Additionally so called cross tree constraints can be defined to model, for example mutual exclusive features across the whole tree structure.\cite{b9}\\

@@

\begin{figure}[h!]
	\includegraphics[width=\columnwidth]{decision_model}
	\caption{Simplified decision model of a car}
\end{figure}
Fig. 2 uses the same model as Fig. 1 but this time with the decision modeling approach was used. As described previously the decision model does not contain information about the commonalities in the system.\\

\section{Variability patterns}
Variability can be introduced into a system via different mechanisms. These mechanisms strive to bind a given system to one out of many alternatives at different binding times. Binding times describe when exactly these mechanisms stop deffering the decision and make the choice for any of the variations, as illustrated in Figure 3.\cite{b2}

@@

Further explanation of these binding times can be found in the subsequent paragraphs:\\
\begin{itemize}
	\item \textit{Product architecture derivation.} The architecture of the product already binds to a variation. This is the lowest binding time and will influence all subsequent steps.\\
	\item \textit{Compilation.} When the source code is compiled there exist several options to bind to a variation by adjusting constants, macros or compiler directives.\\
	\item \textit{Linking.} This option does not exist for any technology, but for many programming languages the linking process can be manually changed. For example by switching the DLL of a module.\\
	\item \textit{Startup} A simple option to bind variants to the system is by configuration files that influence the features a software system exhibits.\\
	\item \textit{Adaptation to runtime environment} This typically refers to plugins of a software, which can add variation to a system and can either be provided by the original developer or even third party vendors. \\
\end{itemize}
Any mechanism to bind to none, one or many different variants tend to fall into different categories presented by \cite{b2} and \cite{b3}, and also have parallels to the feature types presented in the previous section:\\
\begin{itemize}
	\item \textit{Variant Entity.} There exist several options but only one may be bound to at any given time.
	\item \textit{Optional Entity.} An optional entity may or may not be included in the system.
	\item \textit{Multiple Coexisting Entities.} Similar to the variant entity but with the exception that multiple variants can be used.\\
\end{itemize}
These categories tend to be the same for the different binding times, but of course vary greatly in their implementation. While an optional entity during compilation may only be influenced by a compiler flag, it might be a whole plugin from a third party vendor when adapting a given runtime environment. An in depth analysis of these mechanisms can be found in \cite{b2}.
